### 2017-07-25 Meeting Notes.Rmd
added meeting notes

---
title: "Meeting Notes 7/25/17"
author: "Tessa McDonnel"
date: "7/25/2017"
output: html_document
---

#### Weston Role for Collaboration

* can provide data on users (1.3) million who have taken classes
* know common mistakes from users
* scts tailored to student mistakes
* 1,000 universities enrolled
* Professors use it as pre-lab exercise or assign entire courses

#### Perspectives on how it works in classroom

* Weston will try to get someone who has experience... might have to wait till October

#### Sharing platform preference

* LaTeX and GitHub
* Use email to communicate

#### Weston Workflow
* what data camp is 
* look over what's written about the how-to

## integrating LMS

* used a lot by professors and works well