### 2017-08-09 Outline.Rmd
updated outline with Mine's input

---
title: "Outline"
author: "Tessa McDonnel"
date: "7/24/2017"
output: html_document
---

### Abstract

### Introduction
* What is DataCamp?
* increased use of R and R Markdown in classrooms- need more effective way to get students familiar with R
* competing technologies- alternative ways for students to learn R in/out of the classroom with interactive platform
    + Swirl: versatile for instructors but can not monitor student work 
    + TryR: not versatile and can not monitor student work
    + LearnR
* Gamification- increased student focus when points are on the line, must actively engage to do well
* Pedagogical Context
    + How technology is meant to be used - can be used as traditional homework assignment, in the classroom during lab periods, as pre-class assignments for flipped classrooms, even as a  test or assessment tools??
    + Intendend users - novice through advanced programming students will benefit, any college educator can access or contribute
    + Skills or concepts learned - focus is on coding in R to obtain statistical results, with some interpretation
*accessibility - available on tablets/mobile devices

### How to use Data Camp

##### Using Courses
* need a DataCamp account
* Easy for students- don't have to download anything because course is web based 
* Take advantage of free community courses- beginner to advanced programming students
* Premium content free for academia- create a Classroom Group to track student progress
* Monitor your own progress with My Profile

##### Creating Course
* What is needed to get started:  DataCamp account (verified instructor), link to GitHub
* Use the template course files from 'Create a DataCamp Course' dialog
* How to start building (including with collaborators)
    + Course Structure
    + Using Teach Editor (as opposed to GitHub)- can easily edit, preview and save changes
    + Examples of DataCamp code for building different types of modules: normal, plain multiple choice, multiple choice for exercises involving graphs/images
    + Capabilities of SCTs
    + Links to some github repositories: 'datacamp/courses-intro-r' as another example of DataCamp code and 'datacamp/testwhat' to help with submission correctness tests
    + Tailoring pre-existing code to fit specific needs
    + Assign course to Classroom Group- can assign specific chapters/entire course/number of points
* Tips for smoother course development process
    + Using LaTeX math symbols to display greek letters, operators, ect.
    + Need to add 'requirements.r' file with devtools to add packages
    + When using datacamp and github, press 'refresh' when available on the screen

### Integrating with LMS
* Easy to track student progress

### iterative development
* not sure where this fits in the outline - but use metadata from course to iteratively develop labs/activities where students are struggling

### Discussion 
* experience of statistics instructors who have created a DataCamp course and used it in the classroom
* overall pros, some cons- doesn't help students learn R Markdown
* student difficulties/feedback?
* how to get premium content free for teachers