### 2017-09-19 research_slides.tex
started presentation outline

\documentclass{beamer}

\usepackage[utf8]{inputenc}

%Information to be included in the title page:
\title{Creating DataCamp courses}
\subtitle{An interactive teaching tool}
\author{Tessa McDonnel}
\institute{Cal Poly San Luis Obispo}
\date{}

\begin{document}

\frame{\titlepage}

\begin{frame}
\frametitle{questions for Pileggi}

1) Should I be talking about using Git/LaTeX/Mendeley? mainly git

2) Am i focusing on what I actually did (installed software, went throug labs, wrote accompaning datacamp course)? or like what im writing for the article?

\end{frame}

\begin{frame}
\frametitle{Purpose of reasearch project}

This summer I worked with Dr. Pileggi and improved upon the STAT 217 labs in R by creating
a supplimental online DataCamp course that accompanies the labs.

\begin{itemize}
 \item<1-> Technology plays a huge role in statistics, and even introductory statistics courses are utilizing statistical software.
 \item<2-> Non-statistics major students have a lot of trouble with R because they've never seen anything like it.
 \item<3-> The previous STAT 217 pre-labs consisted of a lab manual and video. \textbf{Should i add more about previous pre-labs?}
 \item<4-> My job this summer was to convert the previous lab materials to an interactive online R programming course
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{What is DataCamp?}
\begin{itemize}
 \item<1-> interactive leaning tool; receive instant personalized feedback
 \item<2-> DataCamp is a web-based application, only need a web browser to access courses (no R installation required)- can be very helpful for intro stats 
 \item<3-> Similar to swirl, educators can create their own DataCamp courses or utilize the free existing courses 
 \item<4-> Free for academic groups 
 \item<5-> DataCamp can be linked to learning management systems, easy to track student progress
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Step One}

Work through all of the existing lab assignments, this includes reading lab manuals, watching videos, taking quiz's and completing the actual lab assignment.
\begin{itemize}
 \item<1-> 
 \item<2-> 
 \item<3-> 
 \item<4-> 
\end{itemize}

\end{frame}

\begin{frame}

Slide showing the code for a coding example with a data set and it's corresponding preview

\end{frame}
\end{document}